### 2024-11-26 paper/paper.qmd
1. construct model 2. Introduction of paper 3. remove unused file 4. remove unused package

---
title: "Focasting The Trend of Vehicle Related Crime In Calgary In 2025"
subtitle: "An Analysis of Calgary's Crime Count From 2018 to 2024"
author: 
  - Haobo Ren
thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
date: 23-November-2024
date-format: long
abstract: "Vehicle-related crime is a long-standing challenge to urban security, so accurate prediction is necessary to effectively allocate police resources and develop security policies. This study examines trends in vehicle-related crime in Calgary, Canada, focusing on two types of crime: Theft of vehicle and Theft from vehicle. A Bayesian regression model is fit to historical crime data from January 2018 to October 2024, accounting for seasonal variations (months), long-term trends (years), and differences between crime categories. The usefulness of Bayesian modeling for crime trend forecasting is demonstrated by this analysis, providing law enforcement and policymakers practical insights to foresee and handle long-term and seasonal crime trends."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)

```

# Introduction

Urban crime is an ongoing problem compromising public confidence, economic stability, and safety. Among the several types of crimes, vehicle-related offenses include theft of vehicles and theft from vehicles remain a major issue in urban areas across the country. These crimes disturb people's life but also burden local governments and law enforcement. Over the years, Calgary, Canada, has seen varying patterns in vehicle-related crimes, which raises issues regarding future developments of these trends. Designing effective preventive plans, maximizing resource allocation, and guaranteeing community safety all depend on an awareness of current trends and future crime count prediction. This study forecasts vehicle-related crime patterns in Calgary up until 2025, therefore addressing these issues.

Analyzing and projecting monthly crime counts for theft from vehicles and theft of vehicles in Calgary is the main goal of this study. I created a Bayesian regression model to assess the effects of time (year and month) and crime category on reported crime counts using historical data spanning January 2018 through October 2024. Offering a comprehensive knowledge of how vehicle-related crimes have developed and may continue to alter in the near future, the model catches both long-term patterns and seasonal changes. The research also offers category-specific forecasts, therefore stressing variations in the seasonal and temporal trends of theft from and of vehicles.

The analysis reveals a continued decline in vehicle-related crimes in Calgary, with a mean change of -60.5 each year, and notable mean differences of -454.8 between the two categories. Theft from vehicles is anticipated to show cyclical surges in the summer months, indicating possible opportunistic habits. Conversely, vehicle theft exhibits a more stable decline throughout the course of the year. The data indicate that, although overall vehicle-related crimes are down, seasonal surges in certain categories persist as a problem for policymakers and law enforcement to tackle.

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

## Overview

We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...

Overview text

## Measurement

Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

## Outcome variables

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

## Predictor variables

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References