### 2024-11-26 paper/paper.qmd
Edit data section and references.bib

---
title: "Focasting The Trend of Vehicle Related Crime In Calgary In 2025"
subtitle: "An Analysis of Calgary's Crime Count From 2018 to 2024"
author: 
  - Haobo Ren
thanks: "Code and data are available at: https://github.com/HaoboRrrr/calgary_vehicle_crime_forcast"
date: 23-November-2024
date-format: long
abstract: "Vehicle-related crime is a long-standing challenge to urban security, so accurate prediction is necessary to effectively allocate police resources and develop security policies. This study examines trends in vehicle-related crime in Calgary, Canada, focusing on two types of crime: Theft of vehicle and Theft from vehicle. A Bayesian regression model is fit to historical crime data from January 2018 to October 2024, accounting for seasonal variations (months), long-term trends (years), and differences between crime categories. The usefulness of Bayesian modeling for crime trend forecasting is demonstrated by this analysis, providing law enforcement and policymakers practical insights to foresee and handle long-term and seasonal crime trends."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(here)
library(knitr)

```

# Introduction

Urban crime is an ongoing problem compromising public confidence, economic stability, and safety. Among the several types of crimes, vehicle-related offenses include theft of vehicles and theft from vehicles remain a major issue in urban areas across the country. These crimes disturb people's life but also burden local governments and law enforcement. Over the years, Calgary, Canada, has seen varying patterns in vehicle-related crimes, which raises issues regarding future developments of these trends. Designing effective preventive plans, maximizing resource allocation, and guaranteeing community safety all depend on an awareness of current trends and future crime count prediction. This study forecasts vehicle-related crime patterns in Calgary up until 2025, therefore addressing these issues.

Analyzing and projecting monthly crime counts for theft from vehicles and theft of vehicles in Calgary is the estimand of this study. I created a Bayesian regression model to assess the effects of time (year and month) and crime category on reported crime counts using historical data spanning January 2018 through October 2024. Offering a comprehensive knowledge of how vehicle-related crimes have developed and may continue to alter in the near future, the model catches both long-term patterns and seasonal changes. The research also offers category-specific forecasts, therefore stressing variations in the seasonal and temporal trends of theft from and of vehicles.

The analysis reveals a continued decline in vehicle-related crimes in Calgary, with a mean change of -60.5 each year, and notable mean differences of -454.8 between the two categories. Theft from vehicles is anticipated to show cyclical surges in the summer months, indicating possible opportunistic habits. Conversely, vehicle theft exhibits a more stable decline throughout the course of the year. The data indicate that, although overall vehicle-related crimes are down, seasonal surges in certain categories persist as a problem for policymakers and law enforcement to tackle, for example, the crime count of Theft From Vehicle in August is expected larger than January by 159.4.

The study is significant for its role in proactive crime prevention and resource optimization. This study offers precise and comprehensible forecasts, providing policymakers and law enforcement with evidence-based insights to inform decision-making. This study addresses a significant research gap and enhances public safety in Calgary.

#Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

## Overview

The raw dataset obtained from Open Data Calgary[@OpenCalgary] was recorded and updated monthly by the Calgary Police Service. The data is considered cumulative as late-reported incidents are often received well after an offence has occurred. An incident is either reported just after the crime happened, or reported on the Calgary Police Service[@calgarypolice].

The data analysis and visualization is done in R[@citeR] with the following Packages: tidyverse[@tidy], janitor[@janitor], arrow[@arrow], rstanarm[@rstanarm], ggplot2[@ggplot2], dplyr[@dplyr], here, knitr.

## Measurement

The data reflects reported crime incidents across Calgary, categorized by community, crime type, and temporal details (year and month). Each row in the dataset represents a summary of crime counts for a specific category and community within a given month and year. The Calgary Police Service serves as the primary source, systematically recording incidents reported by the public. These reports may be filed in several ways, including: immediate Reporting and delayed Reporting.

Crime count quantifies the number of reported incidents of a specific crime and time period and is based on the most serious violation per incident. The count represents and aggregation of individual reports collected by the Calgary Police Services.

There are limitations in these measurement. Reporting Bias: Not all crimes are reported. Minor incidents or those involving uninsured vehicles may go unreported, leading to underestimation of true crime rates. Cumulative Nature: The inclusion of late-reported incidents makes it challenging to distinguish between crimes that occurred during the reported month versus those that occurred earlier.

By understanding how this dataset measures real-world crime phenomena, we can more confidently interpret the insights and predictions derived from it.

## Data Examination {#sec-dataexam}

The raw data from Open Data has five columns and 75,595 rows, the column names are displayed below:

```{r}
#| echo: false
#| warning: false
raw_data <- read_csv(here("data/01-raw_data/raw_data.csv"))

# Get the column names of your raw data
column_names <- colnames(raw_data)

# Convert the column names to a data frame for kable
column_df <- as.data.frame(column_names)

# Display with kable
kable(column_df)
```

```{r}

```

Talk way more about it.

## Predictor variables

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

```

\newpage

# References