### 2024-12-04 paper/paper.qmd
Implementation of the survey

---
title: "Focasting The Trend of Vehicle Related Crime In Calgary In 2025"
subtitle: "An Analysis of Calgary's Crime Count From 2018 to 2024"
author: 
  - Haobo Ren
thanks: "Code and data are available at: https://github.com/HaoboRrrr/calgary_vehicle_crime_forcast"
date: 23-November-2024
date-format: long
abstract: "Vehicle-related crime is a long-standing challenge to urban security, so accurate prediction is necessary to effectively allocate police resources and develop security policies. This study examines trends in vehicle-related crime in Calgary, Canada, focusing on two types of crime: Theft of vehicle and Theft from vehicle. A Bayesian regression model is fit to historical crime data from January 2018 to October 2024, accounting for seasonal variations (months), long-term trends (years), and differences between crime categories. The usefulness of Bayesian modeling for crime trend forecasting is demonstrated by this analysis, providing law enforcement and policymakers practical insights to foresee and handle long-term and seasonal crime trends."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(here)
library(knitr)
library(arrow)
library(ggplot2)
library(bayesplot)

cleaned_data <- read_parquet(here("data/02-analysis_data/calgary_crime.parquet"))
fitted_data <- read_parquet(here("data/02-analysis_data/fitted_calgary_crime.parquet"))
raw_data <- read_csv(here("data/01-raw_data/raw_data.csv"))
```

# Introduction

Urban crime is an ongoing problem compromising public confidence, economic stability, and safety[@un2012enhancing]. Among the several types of crimes, vehicle-related offenses include theft of vehicles and theft from vehicles remain a major issue in urban areas across the country[@rakauskas2009identification]. These crimes disturb people's life but also burden local governments and law enforcement. Over the years, Calgary, Canada, has seen varying patterns in vehicle-related crimes, which raises issues regarding future developments of these trends. Designing effective preventive plans, maximizing resource allocation, and guaranteeing community safety all depend on an awareness of current trends and future crime count prediction. This study forecasts vehicle-related crime patterns in Calgary up until 2025, therefore addressing these issues.

Analyzing and projecting monthly crime counts for theft from vehicles and theft of vehicles in Calgary is the main aim of this study. I created a Bayesian regression model to assess the effects of time (year and month) and crime category on reported crime counts using historical data spanning January 2018 through October 2024. Offering a comprehensive knowledge of how vehicle-related crimes have developed and may continue to alter in the near future, the model catches both long-term patterns and seasonal changes. The research also offers category-specific forecasts, therefore stressing variations in the seasonal and temporal trends of theft from and of vehicles.

The analysis reveals a continued decline in vehicle-related crimes in Calgary, with a mean change of -60.5 each year, and notable mean differences of -454.8 between the two categories. Theft from vehicles is anticipated to show cyclical surges in the summer months, indicating possible opportunistic habits. Conversely, vehicle theft exhibits a more stable decline throughout the course of the year. The data indicate that, although overall vehicle-related crimes are down, seasonal surges in certain categories persist as a problem for policymakers and law enforcement to tackle, for example, the crime count of Theft From Vehicle in August is expected larger than January by 159.4. Details are in @sec-result.

The study is significant for its role in proactive crime prevention and resource optimization. This study offers precise and comprehensible forecasts, providing policymakers and law enforcement with evidence-based insights to inform decision-making. This study addresses a significant research gap and enhances public safety in Calgary.

The remainder of this paper is structured as follows: @sec-data provides an overview of the data. @sec-model provides the modeling approach. @sec-result present the results and discuss the implications, limitations.  @sec-discussion.

# Data {#sec-data}

## Overview

The raw dataset obtained from Open Data Calgary[@OpenCalgary] was recorded and updated monthly by the Calgary Police Service. The data is considered cumulative as late-reported incidents are often received well after an offence has occurred. An incident is either reported just after the crime happened, or reported on the Calgary Police Service[@calgarypolice].

The data analysis and visualization is done in R[@citeR] with the following Packages: tidyverse[@tidy], janitor[@janitor], arrow[@arrow], rstanarm[@rstanarm], ggplot2[@ggplot2], dplyr[@dplyr], here[@here], knitr[@knitr], and bayesplot[@bayes].

## Measurement

The data reflects reported crime incidents across Calgary, categorized by community, crime type, and temporal details (year and month). Each row in the dataset represents a summary of crime counts for a specific category and community within a given month and year. The Calgary Police Service serves as the primary source, systematically recording incidents reported by the public. These reports may be filed in several ways, including: immediate Reporting and delayed Reporting.

Crime count quantifies the number of reported incidents of a specific crime and time period and is based on the most serious violation per incident. The count represents and aggregation of individual reports collected by the Calgary Police Services.

There are limitations in these measurement. Reporting Bias: Not all crimes are reported. Minor incidents or those involving uninsured vehicles may go unreported, leading to underestimation of true crime rates. Cumulative Nature: The inclusion of late-reported incidents makes it challenging to distinguish between crimes that occurred during the reported month versus those that occurred earlier.

By understanding how this dataset measures real-world crime phenomena, we can more confidently interpret the insights and predictions derived from it.

## Data Examination {#sec-dataexam}

The raw data from Open Data has five columns and 75,595 rows, the column names are displayed below: @tbl-raw

```{r}
#| label: tbl-raw
#| tbl-cap: Column Names of Raw Data
#| echo: false
#| warning: false

# Get the column names of your raw data
kable(head(raw_data))
```
- Community: The dataset is spatially disaggregated into Calgary's various communities

Numeric Variable:

- Year: The year of each record indicate when the incidents were reported.

Categorical Variable:

- Category: Each entry is categorized into distinct crime types.

- month: The month of each incidents recorded. Range from 1-12 repeatedly each year

Response Variable:

- Crime Count: This variable quantifies the number of reported incidents of a specific crime type (e.g., theft from or of vehicles) within a given community and time period.

## Data Cleaning

Several steps was taken for better analysis of vehicle-related trend in Calgary. First, clean the names of all the columns. Second, the raw dataset contained granular records categorized by Community and specific details. These were aggregated to provide summarized counts of crimes (crime_count) by category, year, and month to align with analysis requirements. Third, a new time column was introduced in the cleaned data, combining year and month into a single date field (YYYY-MM-DD). This was essential for time-series analysis and visualization. Last, filter out the categories related to vehicle.

## Cleaned Data

The first 6 rows of the cleaned dataset are shown in @tbl-cleaned-data.

```{r}
#| label: tbl-cleaned-data
#| tbl-cap: Head of cleaned Calgary Crime data 
#| echo: false
#| warning: false

kable(head(cleaned_data))
```

## Response Variable

### Crime count of Theft FROM Vehicle

Summary statistics of Crime count of Theft FROM Vehicle are shown in the @tbl-response-theftfromvehicle

```{r}
#| label: tbl-response-theftfromvehicle
#| tbl-cap: Summary Statitic of crime_count(Theft FROM Vehicle)
#| echo: false
#| warning: false

kable(cleaned_data %>% filter(category == "Theft FROM Vehicle") %>% select(crime_count) %>% summary(crime_count))
```

The response variable, crime_count, represents the number of reported incidents of theft from vehicles within Calgary. The summary statistics reveal a significant variability in the crime counts, ranging from a minimum of 120 to a maximum of 1,619 reported cases. The median value is 916, indicating that half of the recorded months experienced fewer than 916 thefts, while the other half exceeded this value. The mean crime count, at 932.7, is slightly higher than the median, suggesting the presence of months with particularly high crime counts that may be skewing the average. The interquartile range (IQR), spanning from 788.2 (1st quartile) to 1,105.5 (3rd quartile), highlights the central tendency of crime occurrences, with most months falling within this range. These statistics emphasize the fluctuating nature of vehicle thefts in Calgary and provide a foundational understanding for modeling and forecasting future trends.

### Crime count of Theft OF Vehicle

Summary statistics of Crime count of Theft FROM Vehicle are shown in the @tbl-response-theftofvehicle

```{r}
#| label: tbl-response-theftofvehicle
#| tbl-cap: Summary Statitic of crime_count(Theft OF Vehicle)
#| echo: false
#| warning: false

kable(cleaned_data %>% filter(category == "Theft OF Vehicle") %>% select(crime_count) %>% summary(crime_count))
```

The response variable, crime_count, for Theft OF Vehicle incidents reflects the monthly reported counts of vehicles being stolen in Calgary. The dataset's summary statistics indicate that the crime counts range from a minimum of 146 to a maximum of 651 incidents. The median count is 424.5, suggesting that half of the months have fewer than 425 thefts, while the other half exceed this value. Interestingly, the mean crime count is 422.5, closely aligned with the median, suggesting a relatively symmetric distribution of the data. The interquartile range (IQR) spans from 361.2 (1st quartile) to 469.0 (3rd quartile), indicating that most monthly counts are concentrated within this range. These figures highlight the scale and regularity of vehicle theft in Calgary and form an essential basis for analyzing and forecasting trends in such crimes.

## Predictor Variable

### year

The variable year represents the timeline over which the crime data is recorded, providing insight into how vehicle-related crime counts have changed annually. By analyzing the data across different years, we can observe any long-term trends, such as consistent increases, decreases, or fluctuations in vehicle thefts. This variable helps us understand whether certain years experienced higher crime rates due to factors like economic conditions, population changes, or other external influences.

### month and category

The variables month and category provide a detailed view of how vehicle-related crimes are distributed across time and crime types. Month captures seasonal variations, showing how crime rates fluctuate throughout the year. For example, there might be a tendency for higher crime counts during certain seasons, such as summer or winter, influenced by weather conditions or holiday activities. Category distinguishes between "Theft FROM Vehicle" and "Theft OF Vehicle," allowing us to compare the frequency and patterns of these two crime types. Together, these variables highlight differences in the nature and timing of vehicle-related crimes, offering valuable insights into their temporal and categorical distributions.

## Trend of The Two Crime

The trend of Theft FROM Vehicle and Theft OF Vehicle are shown in @fig-combined-trend

```{r}
#| label: fig-combined-trend
#| fig-cap: Trend of The Two Crime Combined
#| echo: false
#| warning: false
# Plot combined trends
ggplot(cleaned_data, aes(x = time, y = crime_count, color = category)) +
  geom_line() +
  labs(
    title = "Trend of Theft FROM Vehicle and Theft OF Vehicle",
    x = "Time",
    y = "Crime Count",
    color = "Category"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
\newpage

The trends for "Theft FROM Vehicle" and "Theft OF Vehicle" reveal distinct patterns over time. "Theft FROM Vehicle" consistently shows higher crime counts compared to "Theft OF Vehicle," indicating that it is a more prevalent issue. The trend for "Theft FROM Vehicle" demonstrates significant fluctuations, with peaks around 2019 and a general decline afterward, suggesting a reduction in these incidents in more recent years. In contrast, "Theft OF Vehicle" exhibits a relatively stable pattern with smaller variations over time. Although it remains less frequent than "Theft FROM Vehicle," its counts appear to decrease gradually, particularly after 2022. Both categories show declining trends toward the end of the observed period, which could reflect successful crime prevention measures or other external influences.

# Model {#sec-model}

The Bayesian regression model was developed to analyze and forecast the trends of vehicle-related crimes in Calgary, focusing on the categories "Theft FROM Vehicle" and "Theft OF Vehicle." The primary goal of this modeling is to understand the relationship between crime counts and key predictors such as time (year and month) and crime categories, as well as their interactions. By incorporating temporal and categorical variables, the model seeks to capture long-term trends, seasonal patterns, and differences between the two types of crimes. Ultimately, the model provides a foundation for projecting future crime rates, offering valuable insights that can support strategic decision-making and the development of targeted crime prevention initiatives for 2025 and beyond.

Background details and diagnostics are included in Appendix @sec-model-detail.

## Model set-up

The Bayesian regression model is set up to examine the relationship between vehicle-related crime counts and the selected predictors. The response variable, $y_i$, represents the observed crime count for the i-th observation, where each observation corresponds to a specific month, year, and crime category. The predictors include a numeric variable for year, categorical variables for month and category, as well as interaction terms (category:month and category:year) to capture how crime patterns differ between the two crime types over time and across seasons.

The model assumes the form:

$$
y_i = \beta_0 + \beta_1 \text{year}_i + \beta_2 \text{month}_i + \beta_3 \text{category}_i + \beta_4 (\text{category} \times \text{month})_i + \beta_5 (\text{category} \times \text{year})_i + \epsilon_i
$$

Here, $\beta_0$ is the intercept, capturing the baseline level of crime counts. The coefficients $\beta_1$, $\beta_2$, and $\beta_3$ quantify the effects of year, month, and category, respectively, while $\beta_4$ and $\beta_5$ capture the interaction effects. The error term $\epsilon_i$ accounts for the residual variability in crime counts not explained by the predictors, modeled with a Gaussian distribution. This setup allows the model to estimate the relative contributions of each predictor to crime counts and their combined effects on the observed trends.

## Model justification

The primary goal of this model is to forecast trends in vehicle-related crimes in Calgary, focusing on two categories: "Theft FROM Vehicle" and "Theft OF Vehicle." To achieve this, a Bayesian regression model was chosen, as it provides a robust framework for incorporating prior knowledge, handling uncertainty, and capturing complex relationships between predictors.

The selection of predictors for the Bayesian regression model is grounded in their expected relationship with the response variable, crime_count, and their importance in explaining and forecasting trends in vehicle-related crimes. Below is a detailed justification for each predictor. More detailed justification and test of model are in @sec-model-detail

### year(Numeric Predictor)

The year variable is included as a numeric predictor because vehicle-related crimes are likely influenced by long-term trends over time. For instance, crime rates may decrease due to improvements in law enforcement strategies or community awareness campaigns. Conversely, they may increase due to economic downturns or population growth. By treating year as a continuous variable, the model can capture these gradual changes and make accurate predictions for future years, such as 2025.

Relationship to Crime Counts: Crime counts for both categories ("Theft FROM Vehicle" and "Theft OF Vehicle") show clear temporal trends when plotted against year. For example, "Theft FROM Vehicle" demonstrates an overall decline in recent years, whereas "Theft OF Vehicle" shows a more gradual change (@fig-year-trend).

```{r}
#| label: fig-year-trend
#| tbl-cap: Trend of Crime Counts Over Years
#| echo: false
#| warning: false

ggplot(cleaned_data, aes(x = year, y = crime_count, color = category)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    title = "Trend of Crime Counts Over Years",
    x = "Year",
    y = "Crime Count",
    color = "Category"
  ) +
  theme_minimal()

```

### month (Categorical Predictor)

The month variable is treated as a categorical predictor to account for seasonal variations in crime rates. Crime counts often fluctuate across months due to changes in weather, holidays, and other seasonal factors. For example, warmer months may see increased activity that could lead to more vehicle-related crimes, while colder months might see a decline.

Relationship to Crime Counts: Monthly crime counts display seasonal patterns, with some months consistently showing higher crime rates than others. Including month as a categorical variable allows the model to account for these patterns without assuming a linear relationship.

### category (Categorical Predictor)

The category variable differentiates between "Theft FROM Vehicle" and "Theft OF Vehicle," two distinct types of crimes with different characteristics and frequencies. This variable is crucial for explaining the variation in crime counts, as the two categories exhibit different trends over time and across months.

Relationship to Crime Counts: "Theft FROM Vehicle" has consistently higher crime counts than "Theft OF Vehicle." Additionally, the trends for these categories vary significantly over time and seasons, as seen in prior visualizations. By including category as a predictor, the model can capture these differences and provide more accurate predictions.

### Interactions (Category:Month, Category:Year)

Including interaction terms between category and time variables (month and year) allows the model to capture how the seasonal and yearly trends differ for "Theft FROM Vehicle" and "Theft OF Vehicle." For example, one category might experience sharper seasonal peaks or steeper declines over the years compared to the other.

Relationship to Crime Counts: Visualizations of crime counts by year and month, separated by category, show evidence of these interactions. For instance, "Theft FROM Vehicle" exhibits more pronounced seasonal patterns and a sharper decline in recent years compared to "Theft OF Vehicle."

# Results {#sec-result}

## Overview

This study aims to address the research question: What is the forecasted trend for vehicle-related crimes in Calgary in 2025? Using a Bayesian regression model, we analyzed historical crime data from 2018 to 2024 and forecasted crime counts for 2025. The results are presented using summary statistics, regression estimates, and visualizations comparing actual and fitted trends.

## Trends in Crime Counts

The actual and fitted crime trends for both categories from 2018 to 2025 are shown in @fig-forcast-trend. The fitted values closely match the actual values, indicating the model's strong predictive accuracy.

```{r}
#| label: fig-forcast-trend
#| fig-cap: Actual vs. Fitted Crime Trends (2018-2025)
#| echo: false
#| warning: false
ggplot(fitted_data, aes(x = time, y = crime_count, color = category, linetype = source)) +
  geom_line() +
  labs(
    title = "Actual vs. Fitted Crime Trends (2018-2025)",
    x = "Time",
    y = "Crime Count",
    color = "Category",
    linetype = "Source"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Theft FROM Vehicle: The crime counts show seasonal variability, with peaks during certain months (e.g., summer) and an overall declining trend. The model forecasts that this category will continue to decrease in 2025, with lower peak values and reduced seasonal fluctuations compared to earlier years.
Theft OF Vehicle: This category exhibits a more consistent decline in crime counts over time, with fewer pronounced seasonal patterns. The model predicts further reductions in 2025, with crime counts stabilizing at relatively low levels.

## Regression Model Estimates

The regression model estimates, summarized in the @tbl-model-summary below, provide insight into how predictors such as year, month, and category influence vehicle-related crime counts in Calgary. The table includes coefficient estimates and their standard deviation to quantify the effect sizes and uncertainties associated with each predictor.

```{r}
#| label: tbl-model-summary
#| tbl-cap: Regression Estimates for Crime Counts
#| echo: false
#| warning: false
library(modelsummary)
crime_model <- readRDS(here("models/crime_model.rds"))

# Extract the model summary as a matrix
model_summary <- summary(crime_model)

# Convert the summary into a data frame
summary_table <- as.data.frame(model_summary)

# Rename the columns for clarity
colnames(summary_table) <- c("Mean", "SD", "10%", "50%", "90%")

# Round the "Mean" and "SD" columns to 2 decimal places
summary_table$Mean <- round(summary_table$Mean, 1)
summary_table$SD <- round(summary_table$SD, 1)

kable(summary_table %>% select("Mean", "SD"))

```

### Year:

The coefficient for Year is -60.5, indicating that on average, crime counts decrease by approximately 60 incidents per year for the reference category. This finding aligns with the observed overall decline in vehicle-related crime over time.

### Month:

Monthly coefficients capture seasonal variations in crime counts relative to the reference month January. July and August exhibit higher-than-average crime counts, with estimates of 123.8 and 159.4, respectively, reflecting seasonal spikes during colder months. In contrast, December shows a lower-than-average count, with a negative estimate of -134.8, indicating fewer crimes during this period compared to January.

### Category:

The coefficient for Theft OF Vehicle is -454.8, showing that crime counts for "Theft OF Vehicle" are consistently lower than those for "Theft FROM Vehicle." This highlights the fundamental difference in frequency between the two categories. This finding reinforces the visual observation that "Theft FROM Vehicle" consistently exhibits higher counts than "Theft OF Vehicle."

### Interaction Terms (Year × Category):

The interaction term for Year and Theft OF Vehicle has an estimate of 0.0, suggesting that the declining trend in crime counts over time is statistically no difference between "Theft OF Vehicle" and "Theft FROM Vehicle."

## Addressing the Research Question

The model's results indicate that both types of vehicle-related crimes in Calgary are expected to decline in 2025. Key findings include:

- Overall Decline: Both "Theft FROM Vehicle" and "Theft OF Vehicle" are forecasted to show lower crime counts compared to previous years. This decline reflects both long-term trends and recent reductions in crime rates.

- Category Differences: "Theft FROM Vehicle" will remain more frequent than "Theft OF Vehicle," but the gap between the two categories is expected to narrow slightly as the decline in the former is more pronounced.

- Seasonal Patterns: Seasonal fluctuations are expected to persist, especially for "Theft FROM Vehicle," but the peaks will be less pronounced than in earlier years.

These findings suggest that crime prevention strategies targeting vehicle-related crimes have been effective and should continue to focus on mitigating seasonal spikes and addressing high-crime months.

The Bayesian regression model provides strong evidence for declining trends in vehicle-related crimes in Calgary. By 2025, both categories are projected to experience reduced crime rates, with "Theft FROM Vehicle" showing a significant decrease in seasonal peaks and overall counts. These results offer valuable insights for law enforcement and policymakers, supporting the development of targeted strategies to maintain and accelerate these positive trends.

# Discussion {#sec-discussion}

## What Is Done in This Paper?

This paper investigates the trends and forecasts of vehicle-related crimes in Calgary, specifically focusing on the categories Theft FROM Vehicle and Theft OF Vehicle. Using historical data from 2018 to 2024 and a Bayesian regression model, the analysis identifies long-term trends, seasonal patterns, and differences between the two categories of crime. The model incorporates key predictors such as year, month, category, and their interactions to capture both temporal and categorical variations. The paper also provides forecasts for 2025, offering valuable insights into the expected future trends of these crimes.

## What Is Something That We Learn About the World?

One key finding from this study is the consistent decline in vehicle-related crimes over the years. The results suggest that crime prevention measures, technological advancements (such as better vehicle security systems), and targeted policing strategies may have contributed to this downward trend. The model also highlights seasonal patterns, particularly for Theft FROM Vehicle, which tends to peak in the summer months. These findings reinforce the idea that external factors such as weather, holiday travel, and increased outdoor activity can influence crime rates.

## What Is Another Thing That We Learn About the World?

Another important insight is the fundamental difference between the two crime categories. Theft FROM Vehicle is significantly more frequent than Theft OF Vehicle, and its seasonal variability is more pronounced. This suggests that theft from vehicles is opportunistic and influenced by environmental factors, whereas vehicle theft itself may be less seasonal and potentially driven by different motives, such as organized crime or professional theft rings. These differences imply that crime prevention strategies should be tailored to the specific characteristics of each category.

## Evaluating Policy Interventions

A critical question arising from this research is how effective existing policy interventions have been in reducing vehicle-related crimes. While the model identifies a downward trend in both crime categories, it does not directly link these trends to specific policies, such as enhanced policing, public awareness campaigns, or technological improvements like vehicle immobilizers. Future research could assess the causal relationship between these interventions and crime reduction by incorporating data on policy implementation dates and intensity. For example, how did increased police patrols in high-crime areas or the introduction of anti-theft programs impact specific crime patterns? Addressing this question would provide a clearer understanding of which strategies work best and how they can be optimized.

## What Are Some Weaknesses of What Was Done?

While the analysis provides valuable insights, it is not without limitations:

- Data Limitations: The study relies on historical data, which may not fully capture all relevant factors influencing crime trends. For example, socioeconomic conditions, population density, or enforcement policies are not explicitly included as predictors in the model.

- Simplified Temporal Assumptions: The model assumes linear trends for year and constant seasonal effects for month. These assumptions may not capture potential nonlinearities or abrupt changes in crime patterns caused by external events (e.g., the COVID-19 pandemic or policy changes).

- Interaction Complexity: Although the inclusion of interactions (e.g., Month × Category) improves the model, it may still oversimplify some complex relationships, particularly those involving latent or unobserved variables.

- Forecasting Uncertainty: While the Bayesian framework quantifies uncertainty, the forecasts for 2025 remain dependent on the assumption that historical patterns will continue, which may not hold if new or unforeseen factors emerge.

\newpage

\appendix

# Model Diagnostics {#sec-model-detail}

## Overview

To ensure the reliability and validity of the Bayesian regression model, a series of diagnostic checks were performed. These diagnostics evaluate the model's fit, the adequacy of the predictors, and whether the assumptions underlying the Bayesian framework are satisfied. Below is a detailed summary of the diagnostics conducted.

## Residual Analysis

Residuals @fig-residual-plot represent the differences between observed and predicted values and can indicate issues with model fit or violations of assumptions.

```{r}
#| label: fig-residual-plot
#| fig-cap: Residual Plot
#| echo: false
#| warning: false

# Residual plot
residuals <- residuals(crime_model)
plot(residuals, ylab = "Residuals", xlab = "Fitted Values", main = "Residual Plot")
abline(h = 0, col = "red", lty = 2)

```

The residual plot shows no discernible pattern, and residuals should be randomly scattered around zero. Patterns in the residuals (e.g., funnel shapes or trends) may indicate heteroscedasticity or omitted variable bias.

## Q-Q Plot

The Q-Q Plot @fig-qq assesses whether the residuals follow a normal distribution, which is an assumption of the Gaussian family used in the model. Deviations from the line in the Q-Q plot may suggest non-normality.

```{r}
#| label: fig-qq
#| fig-cap: Q-Q plots of Residuals
#| echo: false
#| warning: false
#| 
# Q-Q plot of residuals
qqnorm(residuals, main = "Q-Q Plot of Residuals")
qqline(residuals, col = "red", lty = 2)

```

Residuals should fall approximately along the reference line. Significant deviations (e.g., at the tails) indicate potential violations of the normality assumption. For Bayesian models, slight deviations are tolerable.

## Posterior Predictive Checks
Posterior predictive checks were performed to evaluate how well the model's predictions align with the observed data. The density overlay plot @fig-posterior compares the observed crime counts to the posterior predictive distribution.

```{r}
#| label: fig-posterior
#| fig-cap: Posterior Predictive Check
#| echo: false
#| warning: false

pp_check(crime_model, plotfun = "dens_overlay") +
  labs(title = "Posterior Predictive Check: Density Overlay")

```

The observed data aligns closely with the predictive distributions, indicating that the model captures the overall distribution of crime counts effectively. Any systematic mismatch could suggest model misspecification, but in this case, the results indicate good fit.

# Idealized Methodology for Data Collection, Analysis and Survey{#sec-ideal-meth}

## Overview

The current Calgary crime dataset, while robust in capturing reported crimes, lacks contextual and unreported crime data that could enrich analyses. An idealized methodology involves supplementing this dataset with survey-based data collection to gather information about unreported crimes, public perceptions of safety, and crime-related behaviors. This approach bridges gaps in observational data and offers a more comprehensive view of crime dynamics in Calgary.

Key goals of this methodology include:

- Capturing unreported crimes to reduce underreporting bias.

- Understanding the demographic and socioeconomic contexts of crime victims.

- Providing actionable insights for law enforcement and policymakers.

## Sampling Approach

Target Population:

- The target population includes Calgary residents aged 18 and above, distributed across all neighborhoods.

Stratified Random Sampling:

- Stratification ensures that the sample reflects Calgary’s diverse population, accounting for factors like neighborhood, income levels, age, and gender.

- Neighborhood-level stratification ensures that areas with low reporting rates are adequately represented.

Sample Size:

- A sample of approximately 2,000 respondents is proposed to ensure statistical power and generalizability, accounting for Calgary’s population size and distribution.

Randomized Sampling:

- Respondents within each stratum are selected using random digit dialing for phone surveys or randomized email invitations for online surveys.

## Respondent Recruitment

Recruitment Strategy:

- ultimodal recruitment methods ensure inclusivity and accessibility:
  - Phone Calls: For residents with landlines or mobile numbers.
  - Online Platforms: Using email invitations and social media campaigns.
  - In-Person Outreach: At community centers or local events, especially in neighborhoods with limited internet access.

Incentives:

- Respondents are offered small incentives, such as gift cards or charitable donations on their behalf, to encourage participation.

Inclusivity Measures:

- Surveys are made available in multiple languages, reflecting Calgary’s linguistic diversity.
Accessibility features, such as larger print and audio formats, are provided for individuals with disabilities.

## Data Validation

Verification of Responses:

- Responses are cross-referenced with existing crime statistics where possible (e.g., verifying reported incidents through CPS records) to ensure consistency.

- Duplicate or inconsistent responses are flagged and removed during preprocessing.

Filtering Out Biases:

- Screening questions are included to ensure respondents meet eligibility criteria (e.g., Calgary residency).

- Statistical techniques, such as post-stratification weighting, are applied to align the sample distribution with Calgary’s population demographics.

Handling Missing or Ambiguous Data:

- Missing responses are minimized through mandatory fields for critical questions.
Ambiguous responses are flagged for follow-up where feasible.

## Survey Structure

Sections and Topics:

- Demographics: Age, gender, income, education level, and neighborhood of residence.
Crime Experience: Details about reported and unreported crimes, including the type, timing, and location of incidents.

- Perceptions of Safety: Questions on how safe respondents feel in their neighborhood and during specific times of the day.

- Behavioral Insights: Questions on precautions taken to avoid crime (e.g., locking vehicles, using anti-theft devices).

Question Types:

- Closed-Ended Questions: For quantifiable data (e.g., “Have you ever been a victim of vehicle-related crime?”).

- Likert Scales: To measure perceptions of safety and trust in law enforcement.

- Open-Ended Questions: To allow respondents to share additional insights.

Length and Accessibility:

- The survey is designed to take no more than 15 minutes to complete, balancing depth of information with respondent engagement.

## Implementation of The Survey: Calgary Vehicle-Related Crime and Safety Survey

### Introduction

Hello, and thank you for participating in this survey! Your responses will help us better understand vehicle-related crimes and public safety in Calgary. The survey will take approximately 10–15 minutes to complete. Your answers are anonymous, and all data will be used for research purposes only.

### Section 1: Demographic Information

* What is your age?
  - Under 18
  - 18–24
  - 25–34
  - 35–44
  - 45–54
  - 55–64
  - 65 or older

* What is your gender?
  - Male
  - Female
  - Non-binary
  - Prefer not to say

* What is your highest level of education?
  - Less than high school
  - High school diploma or equivalent
  - Some college/university
  - Bachelor’s degree
  - Graduate degree

* What is your household income?
  - Less than $30,000
  - \$30,000 – \$59,999
  - \$60,000 – \$99,999
  - \$100,000 – \$149,999
  - \$150,000 or more
  - Prefer not to say

* In which neighborhood of Calgary do you reside?
  - (Dropdown menu of Calgary neighborhoods or a text field for manual entry)

### Section 2: Crime Experience

* Have you ever experienced vehicle-related crime in Calgary?
  - Yes
  - No

* (If Yes) What type of vehicle-related crime did you experience? (Check all that apply)
  - Theft FROM Vehicle (e.g., stolen items from inside or outside your car)
  - Theft OF Vehicle (e.g., stolen car)
  - Vandalism (e.g., broken windows, graffiti)

* When did the most recent vehicle-related crime occur?
  - Within the past month
  - Within the past 6 months
  - Within the past year
  - More than a year ago

* Did you report this incident to the Calgary Police Service?
  - Yes
  - No
  - I’m not sure

* (If No) Why didn’t you report the incident? (Check all that apply)
  - It wasn’t serious enough
  - I didn’t think the police could help
  - I didn’t have time to report it
  - I didn’t know how to report it
  - Other (please specify): __________

\newpage

# References